---
title: "Day 1"
author: "Jan Savinc"
date: "19/04/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(tidyverse)
library(data.table)
```

# Load data

```{r}
load(file = "../AALHD_2021/R Syntax and Data/R Data/MBSdata.RData")
load(file = "../AALHD_2021/R Syntax and Data/R Data/PBSdata.RData")
load(file = "../AALHD_2021/R Syntax and Data/R Data/HMDSdata.RData")
load(file = "../AALHD_2021/R Syntax and Data/R Data/Dthdata.RData")

# dispensing <- read_csv(file = "../AALHD_2021/R Syntax and Data/R Data/PBSdata.RData")
```


# Define inclusion codes

```{r}
codes_mbs <- c(66551, 66554, 66557, 73815, 73840, 66319, 66322)

codes_oral_hypoglycaemics <- c(1202, 1801, 2178, 2430, 2440, 2449, 2607, 2720, 2939, 2940, 8188, 8189, 8391, 8392, 8450, 8451, 8452, 8533, 8535, 8607, 8687, 8688, 8689, 8690, 8691, 8692, 8693, 8694, 8695, 8696, 8810, 8811)

codes_insulin <- c(1425, 1426, 1429, 1430, 1431, 1461, 1462, 1531, 1532, 1533, 1534, 1535, 1537, 1591, 1592, 1710, 1711, 1713, 1715, 1716, 1718, 1721, 1722, 1761, 1762, 1763, 2061, 2062, 8006, 8084, 8085, 8212, 8390, 8435, 8571, 8609)


## ICD codes have been modified for regex use
codes_diabetes_mellitus_icd9 <- c("^250\\.", "V77\\.1")
codes_diabetes_mellitus_icd10 <- paste0("^E", 10:14, "\\.")

codes_deaths_diabetes_mellitus_icd9 <- c("^250", "^V771")
codes_deaths_diabetes_mellitus_icd10 <- paste0("^E", 10:14)
```


# Coding individual datasets

## Steps 1-3, 5-7, 9-11, 13-15

```{r}
tag_records <- function(data_tbl, codes_var, target_var, inclusion_codes) {
  codes_var <- enquo(codes_var)
  target_var <- enquo(target_var)
  data_tbl %>%
  mutate(!!target_var := !!codes_var %in% inclusion_codes)
}

tag_records_regex <- function(data_tbl, codes_var, target_var, inclusion_codes) {
  codes_var <- enquo(codes_var)
  target_var <- enquo(target_var)
  inclusion_codes = paste(inclusion_codes, collapse = "|")
  data_tbl %>%
  mutate(!!target_var := str_detect(string = !!codes_var, pattern = inclusion_codes))
}

# regex_hmds <- paste(c(codes_diabetes_mellitus_icd9,codes_diabetes_mellitus_icd10), collapse = "|")

tag_wide_records_regex <- function(data_tbl, cols_regex, target_var, inclusion_codes) {
  target_var <- enquo(target_var)
  inclusion_codes <- paste(inclusion_codes, collapse = "|")  # make a single regex out of multiple codes
  data_tbl %>%
    rowwise() %>%
    mutate(!!target_var := any(grepl(x = c_across(cols = matches(cols_regex)), pattern = inclusion_codes))) %>%
    ungroup
}

HMDSdata %>%
  rowwise() %>%
  mutate(diabetes = any(grepl(x = c_across(cols = matches("^diag")), pattern = regex_hmds)))
  ungroup

sort_by_id_and_date <- function(data_tbl, date_var) {
  date_var <- enquo(date_var)
  data_tbl %>%
    arrange(rootlpno, !!date_var)
}

add_file_index <- function(data_tbl) {
  data_tbl %>%
    mutate(fileseq = 1:n())
}

add_morb_index <- function(data_tbl) {
  data_tbl %>%
    group_by(rootlpno) %>%
    mutate(morbseq = 1:n()) %>%
    ungroup
}

add_service_index <- function(data_tbl, target_var) {
  target_var <- enquo(target_var)
  data_tbl %>%
    group_by(rootlpno) %>%
    mutate(servseq = if_else(condition = !!target_var, true = cumsum(!!target_var), false = NA_integer_)) %>%  # index of target record, within-id
    ungroup
}

MBSdata2 <-
  MBSdata %>%
  as_tibble() %>%
  # sort_by_id_and_date(date_var = servdate) %>%
  tag_records(data_tbl = ., codes_var = mbsitem, target_var = glyhb, inclusion_codes = codes_mbs) %>%
  add_file_index() %>%
  add_morb_index() %>%
  add_service_index(target_var = glyhb)

PBSdata2 <-
  PBSdata %>%
  as_tibble() %>%
  # sort_by_id_and_date(date_var = dispdate) %>%
  tag_records(data_tbl = ., codes_var = pbsitem, target_var = diabmed, inclusion_codes = c(codes_oral_hypoglycaemics, codes_insulin)) %>%
  add_file_index() %>%
  add_morb_index() %>%
  add_service_index(target_var = diabmed)

HMDSdata2 <-
  HMDSdata %>%
  as_tibble() %>%
  # sort_by_id_and_date(date_var = sepdate) %>%
  tag_wide_records_regex(data_tbl = ., cols_regex = "^diag", target_var = diabetes, inclusion_codes = c(codes_diabetes_mellitus_icd9, codes_diabetes_mellitus_icd10)) %>%
  add_file_index() %>%
  add_morb_index() %>%
  add_service_index(target_var = diabetes)

## alrady a "type 1" file, one record per individual
Dthdata2 <-
  Dthdata %>%
  as_tibble() %>%
  tag_records_regex(data_tbl = ., codes_var = codcode, target_var = diabetes, inclusion_codes = c(codes_deaths_diabetes_mellitus_icd9, codes_deaths_diabetes_mellitus_icd10))

## save for later use
save(list = c("MBSdata2", "PBSdata2", "HMDSdata2", "Dthdata2"), file = "./data2_files_day1.RData")
```

## Checks: steps 4, 8, 12, 16

```{r}
stopifnot(MBSdata2 %>% filter(glyhb) %>% nrow == 29571)
stopifnot(MBSdata2 %>% filter(servseq==1) %>% nrow == 7239)

stopifnot(PBSdata2 %>% filter(diabmed) %>% nrow == 174131)
stopifnot(PBSdata2 %>% filter(servseq==1) %>% nrow == 6820)

stopifnot(HMDSdata2 %>% filter(diabetes) %>% nrow == 21274)
stopifnot(HMDSdata2 %>% filter(servseq==1) %>% nrow == 5328)

stopifnot(Dthdata2 %>% filter(diabetes) %>% nrow == 965)
```


# Steps 4,8,12,16: "Cutting down"

```{r}
MBScutdown <-
  MBSdata2 %>%
  filter(glyhb) %>%
  rename(date = servdate, sequence = servseq) %>%
  mutate(type=1) %>%
  select(rootlpno, date, age, sex, sequence, type)

PBScutdown <-
  PBSdata2 %>%
  filter(diabmed) %>%
  rename(date = dispdate, sequence = servseq) %>%
  mutate(type = 2) %>%
  select(rootlpno, date, age, sex, sequence, type)

HMDScutdown <-
  HMDSdata2 %>%
  filter(diabetes) %>%
  rename(date = sepdate, sequence = servseq) %>%
  mutate(type = 3)  %>%
  select(rootlpno, date, age, sex, sequence, type)

Dthcutdown <-
  Dthdata2 %>%
  filter(diabetes) %>%
  rename(date = dthdate) %>%
  mutate(type = 4) %>%
  mutate(sequence = 1) %>%  # set sequence to 1 for every death record
  select(rootlpno, date, age, sex, sequence, type)
```


# Step 17: Vertical merge

```{r}
diabantepenult <-
  bind_rows(
    MBScutdown,
    PBScutdown,
    HMDScutdown,
    Dthcutdown
  ) %>%
  add_file_index() %>%
  sort_by_id_and_date(date_var = date) %>%
  add_morb_index()

stopifnot(nrow(diabantepenult)==225941)
```

# Step 18: "Cut down further" to only those with sequence = 1 (or missing in death files)

```{r}
diabpenult <-
  diabantepenult %>%
  filter(sequence == 1) %>%
  select(-sequence) %>%
  add_file_index() %>%
  sort_by_id_and_date(date_var = date) %>%  # problem! this changes the results because the
  add_morb_index()

stopifnot(nrow(diabpenult)==20352)
```


# Step 19: First occurence / incidence record of diabetes

```{r}
# platform <-
#   diabpenult %>%
#   group_by(rootlpno) %>%
#   slice_min(order_by = date, n = 1) %>%
#   ungroup %>%
#   select(-morbseq) %>%
#   add_file_index()

platform <- 
  diabpenult %>%
  filter(morbseq==1) %>%
  select(-morbseq) %>%
  add_file_index()

# platform %>% group_by(rootlpno, type) %>% filter(n()>1)
platform %>% count(type)


stopifnot(nrow(platform)==10675)
stopifnot(platform %>% filter(type == 1) %>% nrow == 5020)
stopifnot(platform %>% filter(type == 2) %>% nrow == 3200)
stopifnot(platform %>% filter(type == 3) %>% nrow == 2425)
stopifnot(platform %>% filter(type == 4) %>% nrow == 30)
```


# Step 20: 

```{r}
diabantepenult %>%
  mutate(yearinc = lubridate::year(date)) %>%
  filter(morbseq == 1) %>%
  count(yearinc) %>%
  print(n=nrow(.))

diabantepenult %>%
  mutate(yearinc = lubridate::year(date)) %>%
  filter(morbseq == 1) %>%
  ggplot(data = ., aes(x=yearinc)) +
  geom_bar() +
  theme_minimal()
```

# Step 21:

```{r}
library(survival)

diabantepenult2 <-
  diabantepenult %>%
  mutate(
    yearinc = lubridate::year(date),
    # previous = as.integer(morbseq < lead(morbseq, n = 1, default = FALSE))
    previous = as.integer(morbseq > 1),
    revsurti = if_else(
      condition = previous==0,
      true = as.integer(date - lubridate::ymd("1989-12-31"), units = "days"),
      false = as.integer(date - lag(date, n=1), units = "days")
      )
    )

survival_step_21 <- survfit(Surv(ceiling(revsurti/90), previous) ~ 1,
conf.type = 'none', data = diabantepenult2)

plot(survival_step_21, xscale = 1/90)

summary(survival_step_21)
# survival_step_21$cumhaz
```

## Life table

> Examine the life table and note where the hazard rate (slope of the curve) hits zero, which is at 2,160 days (make sure your hazard rates are displayed to 4-5 decimal places, otherwise a zero hazard rate may appear to occur earlier). At this point the survival is only 0.040 (4%). This means that in the absence of any documented history of whether or not an earlier event occurred, there is only a 4% chance that an apparent first-time event is genuine. Compare this to a documented event-free history of just 90 days, where there is a 21% chance that an apparent first-time event is genuine.

```{r}

# install.packages("discSurv")
library(discSurv)  # for lifeTable()  - which doesn't seem to work
library(MortalityTables)  # also for lifeTable()

discsurv_lifetable <-
  diabantepenult %>%
  mutate(
    timecol = as.numeric(revsurti) + 1,
    censcol = as.numeric(previous)
  ) %>%
  as.data.frame() %>%
  discSurv::lifeTable(dataSet = ., timeColumn = "timecol", censColumn = "censcol")

## how do we get to the 4-5 digit 0 rounding?!
discsurv_lifetable$Output %>% filter(hazard < 0.00005)

# MortalityTables::lifeTable()
```

# Step 22

```{r}
diabantepenult3 <-
  diabantepenult2 %>%
  mutate(
    cx = if_else(
      condition = revsurti >= 2160,
      true = 1,
      false = 0.04 / (1.017 * (revsurti^-0.438))
    )
  )

diabantepenult3 %>%  # check corrected backcasting figures
  filter(morbseq==1) %>%
  group_by(yearinc) %>%
  summarise(
    uncorrected = n(),
    corrected = sum(cx)
  )
```

# Step 23

```{r}
diabantepenult23 <-
  diabantepenult3 %>% 
  arrange(desc(fileseq)) %>%
  group_by(rootlpno) %>%  # each individual gets tagged by whether they had records in any of the 4 sources
  mutate(
    mbs = any(type == 1),
    pbs = any(type == 2),
    hmds = any(type == 3),
    dth = any(type == 4),
    cth = any(type %in% c(1,2)),
    state = any(type %in% c(3,4))
  ) %>%
  ungroup %>%
  arrange(fileseq)
```


# Step 24

>where N is the estimated total number of cases, n1 is the marginal total of Commonwealth registrants, n2 is the marginal total of State registrants and n12 is the number of individuals registered in both Commonwealth and State data sets. Here, N = 11,631 compared with the observed total count of 10,675, suggesting an ascertainment fraction of 92%.

```{r}
diabantepenult23 %>%
  filter(morbseq==1) %>%
  count(cth, state)

n1 <- diabantepenult23 %>% filter(morbseq==1 & cth) %>% nrow
n2 <- diabantepenult23 %>% filter(morbseq==1 & state) %>% nrow
n12 <- diabantepenult23 %>% filter(morbseq==1 & state & cth) %>% nrow

big_n <- (n1 + 1) * (n2 + 1) / (n12 + 1)
actual_n <- diabantepenult23 %>% filter(morbseq==1) %>% nrow

ascertainemnt_fraction = actual_n / big_n

ci_hi <- big_n + qnorm(0.975) * sqrt( (n1 + 1) * (n2 + 1) * (n1 - n12) * (n2 - n12) / ((n12 + 1)^2) / (n12 + 2))
ci_lo <- big_n + qnorm(0.025) * sqrt( (n1 + 1) * (n2 + 1) * (n1 - n12) * (n2 - n12) / ((n12 + 1)^2) / (n12 + 2))


diabantepenult23 %>%
  filter(morbseq==1) %>%
  count(dth, hmds, pbs, mbs)

f1 <- flextable::flextable(as.data.frame(xtabs(~ cth + state, data=diabantepenult23 %>% filter(morbseq==1))))
f2 <- flextable::flextable(as.data.frame(xtabs(~ mbs + pbs + hmds + dth, data=diabantepenult23 %>% filter(morbseq==1))))
```



# Save results

```{r}
save.image(file = "./day1.RData")
```

