---
title: "Day 4"
author: "Jan Savinc"
date: "22/04/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(tidyverse)
library(lubridate)
library(tictoc)
library(survival)
library(survminer)  # for ggplot plots of survival
```

# Load data

```{r}
load(file = "../AALHD_2021/R Syntax and Data/R Data/interim tables/HMDSdata3.RData")
load(file = "../AALHD_2021/R Syntax and Data/R Data/interim tables/PBSdata3.RData")
load(file = "../AALHD_2021/R Syntax and Data/R Data/interim tables/diabetes8.RData")

## convert data frames to tibbles for easier looking up while coding
HMDSdata3 <- as_tibble(HMDSdata3)
PBSdata3 <- as_tibble(PBSdata3)
diabetes8 <- as_tibble(diabetes8)
```


# Define inclusion codes

```{r}
codes_hmds_ugc_complications <- c(
  "531.X", "532.X", "533.X", "534.X", "535.X", "578.X",
  "K25.X", "K26.X", "K27.X", "K28.X", "K29.X", "K92.0", "K92.1", "K92.2"
)

code_atc_indomethacin <- "M01AB01"

codes_pbs_indomethacin <- c(2454,2459,2757,5126,5127,5128)

# datapasta::tribble_paste()
codes_tbl_indomethacin <-
  tibble::tribble(
  ~pbsitem,     ~strength, ~quantity,
       2454L,  "25mg   Oral",      100L,
       2459L,  "25mg   Oral",      100L,
       2757L,  "100mg  Supp",       20L,
       5126L,  "25mg   Oral",      100L,
       5127L,  "25mg   Oral",      100L,
       5128L, "100mg   Supp",       20L
  ) %>% extract(col = strength,into = c("dose","form"), regex = "(\\d{2,3})mg\\s+(\\w+)", remove = TRUE) %>% mutate(dose = as.numeric(dose))

make_regex_from_codes <- function(codes) {
  codes %>%
    str_remove_all(string = ., pattern="\\.X$") %>%  # remove final .X wildcard, not needed for regex matching!
    str_replace_all(string = ., pattern = "\\.", replacement = "\\\\.") %>%  # replace dot with string literal dot
    paste0("^", ., collapse = "|")
}


```

# Steps 1-2

```{r}
HMDSdata3 <-
  HMDSdata3 %>%
  mutate(
    ugc = as.integer(str_detect(string = diag1, pattern = make_regex_from_codes(codes_hmds_ugc_complications)))  # principal diagnosis onlu
    ) %>%
  group_by(rootlpno) %>%
  mutate(ugcseq = if_else(condition = ugc==1, true = cumsum(ugc==1), false = NA_integer_)) %>%
  ungroup

index <-
  HMDSdata3 %>%
  filter(ugcseq==1 & admdate >= ymd("1991-01-31")) %>%
  select(rootlpno, age, sex, admdate)

stopifnot(nrow(index)==970)
```

# Step 3

```{r}
indexposure <-
  PBSdata3 %>%
  mutate(indometh = as.integer(pbsitem %in% codes_tbl_indomethacin$pbsitem)) %>%
  filter(indometh==1)

stopifnot(nrow(indexposure)==5351)  
```

# Step 4

```{r}
indexposure2 <-
  indexposure %>%
  left_join(index %>% select(rootlpno, admdate), by="rootlpno") %>%
  filter(!is.na(admdate))  # index cases are ones with complications, and hence an admdate in HMDSdata3 tagged for UGC events
```

# Step 5

```{r}
indexposure3 <-
  indexposure2 %>%
  mutate(
    exposure = case_when(
      as.integer(admdate - dispdate, units = "days") %in% 0:31 ~ 1L,
      as.integer(admdate - dispdate, units = "days") %in% 365:395 ~ 2L,
      TRUE ~ 0L
    )
  )
```

# Step 6

```{r}
indexposure4 <-
  indexposure3 %>%
  filter(exposure %in% 1:2) %>%
  group_by(rootlpno, exposure) %>%
  slice_max(order_by = dispdate, n = 1) %>%
  ungroup %>%
  select(rootlpno, dispdate, exposure)
```

# Step 7

```{r}
indcasexp <- indexposure4 %>% filter(exposure==1)
indconexp <- indexposure4 %>% filter(exposure==2)
```

# Step 8

```{r}
index2 <-
  index %>%
  left_join(indcasexp %>% rename(casexdat = dispdate, casexp = exposure)) %>%
  left_join(indconexp %>% rename(conexdat = dispdate, conexp = exposure)) %>%
  mutate(
    casexp = replace_na(casexp, replace = 0),
    conexp = replace_na(conexp, replace = 0) %>% if_else(.==2, 1, .)
  )
```

# Step 9

```{r}
index2 %>%
  count(casexp, conexp)

odds_step9 <- sum(index2$casexp==1 & index2$conexp==0) / sum(index2$casexp==0 & index2$conexp==1)

# xtabs(.~conexp+casexp, data = index2, )
```

# Step 10

```{r}
set.seed(2.5e6)
indexadm <-
  index2 %>%
  mutate(
    randsort = runif(n = n(), min = 0, max = 1)
    ) %>%
  arrange(randsort) %>%
  mutate(randseq = row_number()) %>%
  select(randseq, admdate)
```

# Steps 11-12

```{r}
set.seed(3.5e6)
reference <-
  diabetes8 %>%
  mutate(
    randsort = runif(n = n(), min = 0, max = 1)
    ) %>%
  arrange(randsort) %>%
  mutate(randseq = rep(1:nrow(indexadm), length.out = n())) %>%
  arrange(randseq) %>%
  left_join(indexadm) %>%
  filter(exit >= admdate) %>%
  select(rootlpno, admdate)
```

# Step 13

```{r}
refexposure <-
  PBSdata3 %>%
  mutate(indometh = as.integer(pbsitem %in% codes_tbl_indomethacin$pbsitem)) %>%
  filter(indometh==1)
```

# Step 14

```{r}
refexposure2 <-
  refexposure %>%
  left_join(reference, by ="rootlpno") %>%
  filter(!is.na(admdate))
```

# Step 15

```{r}
refexposure3 <-
  refexposure2 %>%
  mutate(
    exposure = case_when(
      as.integer(admdate - dispdate, units = "days") %in% 0:31 ~ 1L,
      as.integer(admdate - dispdate, units = "days") %in% 365:395 ~ 2L,
      TRUE ~ 0L
    )
  ) %>%
  filter(exposure %in% 1:2)

# refexposure3 %>% count(exposure)
```

# Step 16

```{r}
refexposure4 <-
  refexposure3 %>%
  group_by(rootlpno, exposure) %>%
  slice_max(order_by = dispdate, n = 1) %>%
  ungroup %>%
  select(rootlpno, dispdate, exposure)
```

# Step 17

```{r}
refcasexp <- refexposure4 %>% filter(exposure==1)
refconexp <- refexposure4 %>% filter(exposure==2)
```

# Step 18

```{r}
reference2 <-
  reference %>%  # reference are ids randomly picked from the diabetes file, independent of whether they had UGC
  left_join(refcasexp %>% rename(casexdat = dispdate, casexp = exposure)) %>%
  left_join(refconexp %>% rename(conexdat = dispdate, conexp = exposure)) %>%
  mutate(
    casexp = replace_na(casexp, replace = 0),
    conexp = replace_na(conexp, replace = 0) %>% if_else(.==2, 1, .)
  )
```

# Step 19

```{r}
reference2 %>% count(casexp, conexp)

odds_step19 <- sum(reference2$casexp==1 & reference2$conexp==0) / sum(reference2$casexp==0 & reference2$conexp==1)
## is this equivalent to running a
# glm(formula = casexp~conexp, family = "binomial", data = reference2)



odds_ratio <- odds_step9 / odds_step19
```

> This is your estimate of the effect of first-time UGC within 31 days of the dispensing date of an indomethacin script compared with at other times, over and above any apparent effect introduced by time-trend bias.

# Overlap between index cases and reference cases?

```{r}
# index2 contains index cases
# reference2 contains reference cases

nrow(index2 %>% inner_join(reference2, by="rootlpno"))
```

# data.table messing around

```{r}
library(data.table)

diabetes9 <- as.data.table(diabetes8)

diabetes9[dead==1 & totpt>500]
diabetes9[,list(dead)]
diabetes9[,dead]
diabetes9[,list(dead,totpt)]
diabetes9[,.(dead,totpt)]  # .() is an alias for list()
diabetes9[,.(d = dead, tpt = totpt)]

diabetes9[, sum((stgpt1+stgpt2)<stgpt3)]
diabetes9[dead==1, sum((stgpt1+stgpt2)<stgpt3)]
diabetes9[dead==0, sum((stgpt1+stgpt2)<stgpt3)]

diabetes9[dead==0, .(stg1and2 = sum(stgpt1+stgpt2))]
diabetes9[dead==0, .(stg1and2 = sum(stgpt1+stgpt2), deads=sum(dead==0), n = .N)]
diabetes9[dead==0, c("stg1and2" = sum(stgpt1+stgpt2), "deads"=sum(dead==0), "n" = .N)]
some_magic_names <- c("stage2","stage3")

diabetes9[dead==0, .(stg1and2 = sum(stgpt1+stgpt2), deads=sum(dead==0), n = .N, ..some_magic_names)]


```

